---
title: "fundsr-intro: Rolling differences and Xetra liquidity plots"
author: "Stanislav Traykov"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fundsr-intro: Rolling differences and Xetra liquidity plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

## Setup
```{r setup}
library(dplyr)
library(ggplot2)
library(lubridate)
library(fundsr)
```

Set up directories.
```{r dirs}
dirs <- c(
    data = file.path("data", "funds"),
    xlm = file.path("data", "xlm"),
    out = "output"
)
for (d in dirs) {
    dir.create(d, recursive = TRUE)
}
```

Populate the XLM directory with some monthly XLM reports
```{r xlm-fetch}
xlm_urls <- c(
    "https://www.cashmarket.deutsche-boerse.com/resource/blob/4844258/91ce589f5309cbbd1ad0c92b3e6cdbda/data/20251130-ETF-ETP-Statistic.xlsx",
    "https://www.cashmarket.deutsche-boerse.com/resource/blob/4795286/394c4451af562507f9def3f39da62242/data/20251031-ETF-ETP-Statistic.xlsx",
    "https://www.cashmarket.deutsche-boerse.com/resource/blob/4725636/2d62a1677b537d996aefc45df3ff21d3/data/20250930-ETF-ETP-Statistic.xlsx",
    "https://www.cashmarket.deutsche-boerse.com/resource/blob/4674370/0de847380722e8e87bc821cb5313ba41/data/20250831-ETF-ETP-Statistic.xlsx"
)
for (url in xlm_urls) {
    fname <- basename(url)
    dest_path <- file.path(dirs[["xlm"]], fname)
    if (!file.exists(dest_path)) {
        download.file(url, destfile = dest_path, mode = "wb", quiet = TRUE)
        Sys.sleep(stats::runif(1, 0.5, 1.0))
    }
}
```

Import all XLM files into a tibble
```{r xlm-read}
if (!exists("xlm_data")) {
    xlm_data <- read_xlm_directory(dirs[["xlm"]])
}
```

Set package options
```{r options}
fundsr_options(
    data_dir = dirs[["data"]],
    out_dir = dirs[["out"]],
    # internal_png = TRUE, # output PNGs without Inkscape
    # inkscape = "path/to/inkscape" # set only if auto-detection fails
)

# Helper to add urls to option fundsr.fund_urls
add_fund_urls(c(
    IUSQ = "https://www.ishares.com/uk/individual/en/products/251850/ishares-msci-acwi-ucits-etf/1535604580409.ajax?fileType=xls&fileName=iShares-MSCI-ACWI-UCITS-ETF-USD-Acc_fund&dataType=fund",
    SPYY = "https://www.ssga.com/ie/en_gb/institutional/library-content/products/fund-data/etfs/emea/navhist-emea-en-spyy-gy.xlsx"
))
```

## Get fund data
Populate funds directory (download files in `fundsr.fund_urls` option)
```{r fund-dl}
download_fund_data()
```

Register data loader
```{r data-loader}
add_data_loader(function() {
    spdr("SPYY", benchmark = "ACWI") # automatically attempts to read <ticker>.xls[x]
    ishs("IUSQ", benchmark = "ACWI", retrieve_benchmark = T) # also retrieve ACWI from file
})
```

Get fund and index data into a big tibble
```{r run-data-loaders}
series <- build_all_series() %>%
    filter(date >= as_date("2012-12-29"))
```

Check contents
```{r contents-check}
series %>% filter(date >= as_date("2015-04-03"))
get_fund_index_map()
```

## Calculate CAGR & log diffs
```{r roll-diffs}
nd <- 365
diffs <- roll_diffs(series, nd, get_fund_index_map())
```

## Plot specs
```{r plot-specs}
no_filter <- NULL
zoom_filter <- function(x) {x %>% filter(date >= as_date("2022-01-01"))}
acwi_funds <- c("spyy", "iusq")
gg_par <- scale_color_manual(values = c("iusq" = "red", "spyy" = "blue"), labels = toupper)

plot_spec <- tribble(
    ~plot_id, ~title, ~data_filter,
    ~gg_params, ~width,  ~height,
    ~funds,

    "ACWI", "SPYY & IUSQ", no_filter,
    gg_par, 14, 9,
    acwi_funds,

    "ACWIz",
    c(en = "SPYY & IUSQ: recent years",
      bg = "SPYY & IUSQ: последни години"),
    zoom_filter,
    gg_par, 14, 9,
    acwi_funds
)
```

## Run!
Run the plots! This outputs SVG files and queues plots for optional PNG export using Inkscape (see blow). It may also output lower-quality PNGs (if option `fundsr.internal_png` is `TRUE`).
```{r run-plots}
p <- run_plots(diffs, nd, plot_spec, xlm_data)
```

## Output
```{r acwi-plots, fig.width=12, fig.height=8, out.width="93%", fig.align="center"}
p[["ACWI"]]
p[["ACWI_L"]]
p[["ACWIz_L"]]
p[["xlm_ACWI"]]
```

Corresponding SVG files should be in the `output` directory.

## Plot in another language
```{r acwi-bg-plots, fig.width=12, fig.height=8, out.width="93%", fig.align="center"}
Sys.setlocale("LC_MESSAGES", "bg_BG.UTF-8") # needed on some systems
Sys.setLanguage("bg")
plot_spec_bg <- plot_spec %>%
    mutate(plot_id = paste0(plot_id, "_bg"))
bg_p <- run_plots(diffs, nd, plot_spec_bg, xlm_data)
bg_p[["ACWIz_bg_L"]]
bg_p[["xlm_ACWI_bg"]]
```

## Optional: higher-quality PNG export
```{r inkscape, eval=FALSE}
export_pngs()
#> Executing "C:/Program Files/Inkscape/bin/inkscape.exe" --actions="export-background:white;file-open:output/ACWI.svg;export-filename:output/ACWI.png;export-width:1300;export-do;file-close;file-open:output/xlm_ACWI.svg;export-filename:output/xlm_ACWI.png;export-width:1300;export-do;file-close;file-open:output/ACWI_L.svg;export-filename:output/ACWI_L.png;export-width:1300;export-do;file-close;file-open:output/ACWIz.svg;export-filename:output/ACWIz.png;export-width:1300;export-do;file-close;file-open:output/ACWIz_L.svg;export-filename:output/ACWIz_L.png;export-width:1300;export-do;file-close;file-open:output/ACWI_bg.svg;export-filename:output/ACWI_bg.png;export-width:1300;export-do;file-close;file-open:output/xlm_ACWI_bg.svg;export-filename:output/xlm_ACWI_bg.png;export-width:1300;export-do;file-close;file-open:output/ACWI_bg_L.svg;export-filename:output/ACWI_bg_L.png;export-width:1300;export-do;file-close;file-open:output/ACWIz_bg.svg;export-filename:output/ACWIz_bg.png;export-width:1300;export-do;file-close;file-open:output/ACWIz_bg_L.svg;export-filename:output/ACWIz_bg_L.png;export-width:1300;export-do;file-close"
#> [1] 0
```
